From physical principles to Hamiltonian mechanics

### 2014-06-28 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Reviewing label names

\begin{defn}\label{combine label}
Let $I_1$ and $I_2$ be two sets of labels. We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map $f:I \rightarrowtail I_1, I_2$ exists that maps an element of $I$ to a pair of elements of $I_1$ and $I_2$.
\end{defn}

\begin{defn}\label{orth}
The two label sets $I_1$ and $I_2$ are said to be \emph{orthogonal} iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
\end{defn}

@@

\begin{defn}\label{statedef}
Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}_i$ a \emph{state}.
\end{defn}

\begin{cor}\label{statedistr}
Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$. In physics term, the distribution can be visualized as a histogram over the discrete labels.
\end{cor}

@@

\begin{prop}\label{detMap}
Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
\end{prop}

\begin{cor}\label{discreteEv}
The evolution of a classical configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
\end{cor}

@@

\begin{defn}\label{disclabelsoverr}
Consider a continuous numeric range. We divide the full range into contiguous cells. Let $I$ be the set of cells. For each cell we have a center value $x: I \mapsto \mathbb{R}$ and a width $w: I \mapsto \mathbb{R}$. $I$ is a set of \emph{discrete labels over a continuous range}.
\end{defn}

\begin{prop}\label{disclabelsoverrdist}
Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell. In physics terms, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of each bin.
\end{prop}

\begin{cor}\label{discreteEv}
Let $f: I \leftrightarrow I$ be a bijective map. We have $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
\end{cor}

@@

\begin{defn}\label{discreteHomogeneous}
A set $I$ of discrete labels over a continuous range is said to be \emph{homogeneous} if $w(i)=k$: the bins are of equal width.
\end{defn}

@@

$\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states. So they are not necessarily orthogonal in the extended phase space. Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined. On that plane we can properly count states and define our invariant.

@@

Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$. If we set $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\cite{Synge}\cite{Lanczos}\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary. In fact, it may not need be constant within and across trajectories.}

>>>>

\begin{defn}\label{labelsCombine}
Let $I_1$ and $I_2$ be two sets of labels. We call $I = \langle I_1, I_2 \rangle$ a \emph{combined set of labels} if an injective map $f:I \rightarrowtail I_1, I_2$ exists that maps an element of $I$ to a pair of elements of $I_1$ and $I_2$.
\end{defn}

\begin{defn}\label{labelsOrth}
The two label sets $I_1$ and $I_2$ are said to be \emph{orthogonal} iff $I = \langle I_1, I_2 \rangle = I_1 \times I_2$.
\end{defn}

@@

\begin{defn}\label{classicalPhaseSpace}
Let $\mathbbm{S}_I$ be the set of all possible configuration states of the infinitesimal subdivision. We call this set \emph{phase space}. We call each element $\mathbbm{s}_i$ a \emph{state}.
\end{defn}

\begin{cor}\label{classicalDistribution}
Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ measures how much of the system can be found in each $\mathbbm{s}_i$. In physics term, the distribution can be visualized as a histogram over the discrete labels.
\end{cor}

@@

\begin{prop}\label{detrevMap}
Let $I$ be a set of labels of a system that undergoes deterministic and reversible evolution. There exists a bijective map $f:I \leftrightarrow I$.
\end{prop}

\begin{cor}\label{detrevDist}
The evolution of a classical configuration state $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i$ under a bijective map is given by $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_{i}=\sum D(f^{-1}(i)) \mathbbm{s}_{i}$.
\end{cor}

@@

\begin{defn}\label{discreteLabelDef}
Consider a continuous numeric range. We divide the full range into contiguous cells. Let $I$ be the set of cells. For each cell we have a center value $x: I \mapsto \mathbb{R}$ and a width $w: I \mapsto \mathbb{R}$. $I$ is a set of \emph{discrete labels over a continuous range}.
\end{defn}

\begin{prop}\label{discreteLabelDist}
Each classical configuration state $\mathbbm{c}_j \in \mathbbm{C}_J$ can be identified by a distribution over states: $\mathbbm{c}_j=\sum D(i) \mathbbm{s}_i=\sum \rho(i) w(i) \mathbbm{s}_i$, where $D:I\rightarrow\mathbb{R}$ is defined as before, and $\rho(i)\equiv D(i) / w(i)$ is the density of the distribution for the cell. In physics terms, the distribution can be visualized as a histogram where $w(i)$, $\rho(i)$ and $D(i)$ are respectively the width, height and area of each bin.
\end{prop}

\begin{cor}\label{discreteLabelEv}
Let $f: I \leftrightarrow I$ be a bijective map. We have $\mathbbm{c}_{j'}=\sum D'(i) \mathbbm{s}_i=\sum \rho'(i) w(i) \mathbbm{s}_i = \sum D(f^{-1}(i)) \mathbbm{s}_i = \sum \rho(f^{-1}(i)) w(f^{-1}(i)) \mathbbm{s}_i$. $\rho'(i) = \rho(f^{-1}(i)) w(f^{-1}(i)) / w(i)$.
\end{cor}

@@

\begin{defn}\label{discreteLabelHomogeneous}
A set $I$ of discrete labels over a continuous range is said to be \emph{homogeneous} if $w(i)=k$: the bins are of equal width.
\end{defn}

@@

$\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states. So they are not necessarily orthogonal in the extended phase space. Looking back at \ref{discreteLabelDef}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) change: this is not the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge de$ is defined. On that plane we can properly count states and define our invariant.

@@

Take the metric from \ref{tdofInvariant}, add multiple independent d.o.f as in \ref{mdofInvariant}, use \ref{genAntisim} with the parameter $s$ instead of $t$ and generator $\mathcal{H}$ instead of $H$. If we set $q^{n+1}=t$ and $p^{n+1}=-e$, we recognise Hamilton equations in the extended phase space.\cite{Synge}\cite{Lanczos}\footnote{As in Struckmeier\cite{Struckmeier}, $d_{s}t$ need not be unitary.}
